---
title: "05cleaning"
output: html_document
date: "2023-07-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Clean the HTML outputs

Some pages have some HTML which needs to be removed because it is being rendered as paragraph text: `<p>&lt;!DOCTYPE html&gt;</p>`.

```{r list html files}
#get the names of all the html files
htmlfiles <- list.files("site/_site")
htmlfiles[4]
#read in the first one
testfile <- readr::read_lines(paste0("site/_site/",htmlfiles[4]))
#create an empty list
tfvec <- c()
#loop through all the lines
for(i in testfile){
  #check if the line matches the string
  tfmatch <- i == "<p>&lt;!DOCTYPE html&gt;</p>"
  #store the True/False value in a vector
  tfvec <- c(tfvec,tfmatch)
}
#find the index of the line with that text
doctypeline <- which(tfvec)
print(doctypeline)
#show the  line
#testfile[174]
testfile[doctypeline]
#replace it
testfile[doctypeline] <- ""
testfile[doctypeline]
#save it as a HTML file to check
write(x = testfile, file=paste0("site/_site/","testfile.html"))
#remove the variable
rm(testfile)
```

Some exploration of other files finds that this code sometimes appears on line 167, and sometimes not at all. 

Can we do this by just testing each line?

```{r create cleaning function for DOCTYPE}
#define a function which takes one argument: htmlfile
cleandoctype <- function(htmlfile){
  #read in that file
  thisfile <- readr::read_lines(paste0(htmlfile))
  #store the number of lines
  filelen <- length(thisfile)
 #loop through all the lines after the first one
    for(linenum in seq(2,length(thisfile))){
      #print(thisfile[linenum])
      if (thisfile[linenum] == "<p>&lt;!DOCTYPE html&gt;</p>"){
        print("HIT!")
        print(linenum)
        #replace specified string
        thisfile[linenum] <- ""
      }
    }
  #return the cleaned version to whatever called this function
  thisfile
  }
```

```{r test function}
#run function on one of the files and store in tempfile
tempfile <- cleandoctype(paste0("site/_site/",htmlfiles[10]))
#write the results over the original
write(x = tempfile, file=paste0("site/_site/",htmlfiles[10]))
```

```{r run function on all files}
#loop through the files
for (i in htmlfiles){
  print(i)
  #extract the last 5 chars
  filetype <- substring(i,nchar(i)-4,nchar(i))
  #check if they end in .html
  ishtml <- filetype == ".html"
  print(filetype)
  #this should be TRUE or FALSE
  #print(ishtml)
  #if it's a html file
  if(ishtml){
    #run function on one of the files and store in tempfile
    tempfile <- cleandoctype(paste0("site/_site/",i))
    #write the results over the original
    write(x = tempfile, file=paste0("site/_site/",i))
  }
}
```

## NOTE: Shorter files/errors

In a previous approach we found some HTML files which had much fewer lines than the rest and were missing the DOCTYPE line and menu (you can also spot these by sorting the files by size in the file explorer).

What all appear to have in common is an apostrophe: Peter's, Guy's, King's (and King's Lynn), George's.

We therefore went back to the rendering markdown notebook and added extra code which use str_replace to replace the apostrophes with an empty string. This generated files which were no longer of different length or size, and when the code in this notebook was used, the pages now included a working menu.


## Fix the menu

We need to ensure this navbar works.

The problem is caused by a duplicate `<html>` and `<head>` tag. 

This and other duplicate lines of code run from lines 634-770. But some of that code helps the table to work, so removing it all would cause new problems.

We *can* just remove `<html>` and `<head>` and the closing tags for those.

`<html>` and `<head>` are on 634 and 635 while `</head>` and `<body>` are on line 769 and 770. There's also an extra `</body>` and `</html>` on lines 864-5.

From 636-642 are some meta tags and a title tag we can also remove. 

From 650-658 is a style tag. Removing this doesn't affect the functionality of the page either.

From 675-683 is another which also can be removed without affecting functionality.

But we leave those lines as it's not worth the time of writing code to remove them.

We can make the dropdowns functional by adding this code, but it breaks their appearance. 


```
<!-- CSS file -->
<link rel="stylesheet" href="https://stackpath.bootstrapcdn.com/bootstrap/4.5.0/css/bootstrap.min.css">

<!-- JavaScript files -->
<script src="https://stackpath.bootstrapcdn.com/bootstrap/4.5.0/js/bootstrap.min.js"></script>
<script src="https://stackpath.bootstrapcdn.com/bootstrap/4.5.0/js/bootstrap.bundle.min.js"></script>
```

Some trial and error (trying each line on its own) helps us fix it: the 4.5 JS file is the one that fixes the behaviour of the dropdown, while the CSS file is the one that breaks the appearance. 

So we just need to replace the JS link. This is on lines 20 and 647.

We create a very similar function to before.

```{r create function to clean JS}
#define a function which takes one argument: htmlfile
cleanbootstrap <- function(htmlfile){
  #read in that file
  thisfile <- readr::read_lines(paste0(htmlfile))
  #store the number of lines
  filelen <- length(thisfile)
 #loop through all the lines after the first one
    for(linenum in seq(2,length(thisfile))){
      #print(thisfile[linenum])
      if (thisfile[linenum] == '<script src="libs/bootstrap-3.3.5/js/bootstrap.min.js"></script>'){
        print("HIT!")
        print(linenum)
        #replace specified string
        thisfile[linenum] <- '<script src="https://stackpath.bootstrapcdn.com/bootstrap/4.5.0/js/bootstrap.min.js"></script>'
      }
    }
  #return the cleaned version to whatever called this function
  thisfile
  }
```

```{r test JS function}
#run function on one of the files and store in tempfile
tempfile <- cleanbootstrap(paste0("site/_site/",htmlfiles[10]))
#write the results over the original
write(x = tempfile, file=paste0("site/_site/",htmlfiles[10]))
```

```{r run JS function on all files}
#loop through the files
for (i in htmlfiles){
  print(i)
  #extract the last 5 chars
  filetype <- substring(i,nchar(i)-4,nchar(i))
  #check if they end in .html
  ishtml <- filetype == ".html"
  print(filetype)
  #this should be TRUE or FALSE
  #print(ishtml)
  #if it's a html file
  if(ishtml){
    #run function on one of the files and store in tempfile
    tempfile <- cleanbootstrap(paste0("site/_site/",i))
    #write the results over the original
    write(x = tempfile, file=paste0("site/_site/",i))
  }
}
```

## Splitting the menu

We have too many options on the menu - it runs off the page. So we need to split it into multiple menus.

The menu begins with the HTML code `<ul class="dropdown-menu" role="menu">` on line 251 or thereabouts.

This problem was tackled before at https://github.com/BBC-Data-Unit/child-speech/blob/main/parameterisation/03renderandclean.Rmd

We need to split the menu first at     

```
<li>
      <a href="DARTFORD-AND-GRAVESHAM-NHS-TRUST.html">DARTFORD-AND-GRAVESHAM-NHS-TRUST</a>'
```

*(Note: previously we split it lower but testing revealed that smaller screen resolutions would need it split higher)*

Again let's create a function that will do this for us. 

```{r create function for splitting the navbar}
#define a function which takes one argument: htmlfile
splitmenu <- function(htmlfile, matchstring,menuletters){
  #read in that file
  thisfile <- readr::read_lines(paste0(htmlfile))
  #store the number of lines
  filelen <- length(thisfile)
 #loop through all the lines after the first one
    for(linenum in seq(2,length(thisfile))){
      #print(thisfile[linenum])
      if (thisfile[linenum] == matchstring){
        print("HIT!")
        print(linenum)
        #replace the line before this one with
        thisfile[linenum-1] <- paste0('</ul></li></ul><ul class="nav navbar-nav navbar-right"></ul><ul class="nav navbar-nav"><li class="dropdown"><a href="#" class="dropdown-toggle" data-toggle="dropdown" role="button" aria-expanded="false">',menuletters,'<span class="caret"></span></a><ul class="dropdown-menu" role="menu"><li>')
      }
    }
  #return the cleaned version to whatever called this function
  thisfile
  }
```

```{r split the menu test}
#store the first split
firstsplit <- '      <a href="DARTFORD-AND-GRAVESHAM-NHS-TRUST.html">DARTFORD-AND-GRAVESHAM-NHS-TRUST</a>'
#store the menu letters
firstletters <- 'D-G'
#run the function on that file with that string
tempfile <- splitmenu(paste0("site/_site/",htmlfiles[22]),
                      firstsplit,
                      firstletters)

#repeat for the next split
secondsplit <- '      <a href="HAMPSHIRE-HOSPITALS-NHS-FOUNDATION-TRUST.html">HAMPSHIRE-HOSPITALS-NHS-FOUNDATION-TRUST</a>'

#store the menu letters
secondletters <- 'H-L'
#run the function on that file with that string
tempfile <- splitmenu(paste0("site/_site/",htmlfiles[22]),
                      secondsplit,
                      secondletters)

#repeat for the next split
thirdsplit <- '      <a href="MAIDSTONE-AND-TUNBRIDGE-WELLS-NHS-TRUST.html">MAIDSTONE-AND-TUNBRIDGE-WELLS-NHS-TRUST</a>'
#store the menu letters
thirdletters <- 'M-P'
#run the function on that file with that string
tempfile <- splitmenu(paste0("site/_site/",htmlfiles[22]),
                      thirdsplit,
                      thirdletters)

#and the next split
fourthsplit <- '      <a href="QUEEN-ELIZABETH-HOSPITAL,-KINGS-LYNN,-NHS-FOUNDATION-TRUST.html">QUEEN-ELIZABETH-HOSPITAL,-KINGS-LYNN,-NHS-FOUNDATION-TRUST</a>'
#store the menu letters
fourthletters <- 'Q-SO'
#run the function on that file with that string
tempfile <- splitmenu(paste0("site/_site/",htmlfiles[22]),
                      fourthsplit,
                      fourthletters)

#and the 5th split
fifthsplit <- '      <a href="ST-GEORGES-UNIVERSITY-HOSPITALS-NHS-FOUNDATION-TRUST.html">ST-GEORGES-UNIVERSITY-HOSPITALS-NHS-FOUNDATION-TRUST</a>'
#store the menu letters
fifthletters <- 'ST-T'
#run the function on that file with that string
tempfile <- splitmenu(paste0("site/_site/",htmlfiles[22]),
                      fifthsplit,
                      fifthletters)


#and the 6th split
sixthsplit <- '      <a href="UNITED-LINCOLNSHIRE-HOSPITALS-NHS-TRUST.html">UNITED-LINCOLNSHIRE-HOSPITALS-NHS-TRUST</a>'
#store the menu letters
sixthletters <- 'U'
#run the function on that file with that string
tempfile <- splitmenu(paste0("site/_site/",htmlfiles[22]),
                      sixthsplit,
                      sixthletters)


#and the last split
lastsplit <- '      <a href="WALSALL-HEALTHCARE-NHS-TRUST.html">WALSALL-HEALTHCARE-NHS-TRUST</a>'
#store the menu letters
lastletters <- 'W-Z'
#run the function on that file with that string
tempfile <- splitmenu(paste0("site/_site/",htmlfiles[22]),
                      lastsplit,
                      lastletters)

#write the results over the original
write(x = tempfile, file=paste0("site/_site/",htmlfiles[22]))
```

This next bit could be more elegant - we need to rethink the function so it takes the actual file rather than the filename, as at the moment we need to write it 7 times in order to run the function 7 times. 

```{r apply to all files}
#loop through the files
for (i in htmlfiles){
  print(i)
  #extract the last 5 chars
  filetype <- substring(i,nchar(i)-4,nchar(i))
  #check if they end in .html
  ishtml <- filetype == ".html"
  print(filetype)
  #this should be TRUE or FALSE
  #print(ishtml)
  #if it's a html file
  if(ishtml){
    
    #run function on one of the files and store in tempfile
    tempfile <- splitmenu(paste0("site/_site/",i),
                      firstsplit,
                      firstletters)
    #write the results over the original
    write(x = tempfile, file=paste0("site/_site/",i))
    
    #run function on one of the files and store in tempfile
    tempfile <- splitmenu(paste0("site/_site/",i),
                      secondsplit,
                      secondletters)
    #write the results over the original
    write(x = tempfile, file=paste0("site/_site/",i))
    
    #run function on one of the files and store in tempfile
    tempfile <- splitmenu(paste0("site/_site/",i),
                      thirdsplit,
                      thirdletters)
    #write the results over the original
    write(x = tempfile, file=paste0("site/_site/",i))
    
    #run function on one of the files and store in tempfile
    tempfile <- splitmenu(paste0("site/_site/",i),
                      fourthsplit,
                      fourthletters)
    #write the results over the original
    write(x = tempfile, file=paste0("site/_site/",i))
    
    #run function on one of the files and store in tempfile
    tempfile <- splitmenu(paste0("site/_site/",i),
                      fifthsplit,
                      fifthletters)
    #write the results over the original
    write(x = tempfile, file=paste0("site/_site/",i))
    
    #run function on one of the files and store in tempfile
    tempfile <- splitmenu(paste0("site/_site/",i),
                      sixthsplit,
                      sixthletters)
    #write the results over the original
    write(x = tempfile, file=paste0("site/_site/",i))
    
    #run function on one of the files and store in tempfile
    tempfile <- splitmenu(paste0("site/_site/",i),
                      lastsplit,
                      lastletters)
    #write the results over the original
    write(x = tempfile, file=paste0("site/_site/",i))
  }
}
```


## Changing the 'Trusts' menu to 'Trusts A-C'

We also need to change the first menu so it now reads A-X. The HTML in question is on lines 246-7:

```
  <a href="#" class="dropdown-toggle" data-toggle="dropdown" role="button" data-bs-toggle="dropdown" aria-expanded="false">
    Trusts
```

We can create a similar function to before - the main difference being that it changes the line itself rather than the one before. 


```{r create function for changing menu name}
#define a function which takes one argument: htmlfile
trustmenuAD <- function(htmlfile, matchstring,menuletters){
  #read in that file
  thisfile <- readr::read_lines(paste0(htmlfile))
  #store the number of lines
  filelen <- length(thisfile)
 #loop through all the lines after the first one
    for(linenum in seq(2,length(thisfile))){
      #print(thisfile[linenum])
      if (thisfile[linenum] == matchstring){
        print("HIT!")
        print(linenum)
        #replace the line with
        thisfile[linenum] <- paste0(matchstring,menuletters)
      }
    }
  #return the cleaned version to whatever called this function
  thisfile
  }
```

Now we test that on one file.

```{r test menu function}
#store the first split
firstsplit <- '    Trusts'
#store the menu letters
firstletters <- ' A-C'
#run the function on that file with that string
tempfile <- trustmenuAD(paste0("site/_site/",htmlfiles[21]),
                      firstsplit,
                      firstletters)
#write the results over the original
write(x = tempfile, file=paste0("site/_site/",htmlfiles[21]))
```

And once we're happy, run it on all .html files in the folder.

```{r apply trustmenuAD to all files}
#loop through the files
for (i in htmlfiles){
  print(i)
  #extract the last 5 chars
  filetype <- substring(i,nchar(i)-4,nchar(i))
  #check if they end in .html
  ishtml <- filetype == ".html"
  print(filetype)
  #this should be TRUE or FALSE
  #print(ishtml)
  #if it's a html file
  if(ishtml){
    
    #run function on one of the files and store in tempfile
    tempfile <- trustmenuAD(paste0("site/_site/",i),
                      firstsplit,
                      firstletters)
    #write the results over the original
    write(x = tempfile, file=paste0("site/_site/",i))
  }
}
```

## Add the menu to the index.html file 

Now all the trust pages are fixed, but our index page doesn't have the menu across the top. 

We can fix that by copying the HTML for the menu across.

First we need to render this, and then move it.

```{r render and move index}
#store the location of the template 
rmarkdown::render("index.Rmd")
#move the file
file.copy(from = "index.html",
          to   = "site/_site/index.html")
```

Note: this doesn't always work, so you can manually render, then copy and paste the new file across to the _site folder.

Now let's copy some HTML from one of the other files.

```{r copy across navbar html to index}
#read in a HTML file
parampage <- readr::read_lines(paste0("site/_site/",htmlfiles[30]))
#store the string we'll look for
findstring <- '<div id="header">'

#store the number of lines
filelen <- length(parampage)
#loop through all the lines after the first one
for(linenum in seq(2,length(parampage))){
  #if the line matches our target, store it
  if (parampage[linenum] == findstring){
    print("HIT!")
    print(linenum)
    #store the line number - this will end up being the last one matched
    lastlinenum <- linenum
    }
}

#loop from that last matching line onwards
for (linenum in seq(lastlinenum,length(parampage))){
  #delete that line (replace with empty string)
  parampage[linenum] <- ""
}

#now we need to grab the lines from the index.html page
#read in a HTML file
indexpage <- readr::read_lines("site/_site/index.html")

#store the number of lines
filelen <- length(indexpage)
#loop through all the lines after the first one
for(linenum in seq(2,length(indexpage))){
  #if the line matches our target, store it
  if (indexpage[linenum] == findstring){
    print("HIT!")
    print(linenum)
    #store the line number - this will end up being the last one matched
    lastindexmatch <- linenum
    }
}

#we know we will need to loop the number of lines we need to transfer across
#so calculate how many lines that is 
#by subtracting the matched line number from the length of the doc
totallines <- length(indexpage) - lastindexmatch 
#loop through the numbers 0 to that number
for (i in seq(0,totallines)){
  #add that number to the respective line numbers in each doc
  copylinenum <- lastindexmatch+i
  pastelinenum <- lastlinenum+i
  parampage[pastelinenum] <- indexpage[copylinenum]
}

#write the results over the original
write(x = parampage, file="site/_site/index.html")
```
### Add the chart to the index

We can also embed a chart into the same file using the same approach. 

To help the code find the positions in the text where it needs to place the charts, we've put two periods in those places in the index.Rmd file. This is rendered as a paragraph with two periods, like so: `<p>..</p>`.

The code below goes through each line of the rendered index.html file and stores the location of those placeholders.


```{r find placemarkers}
#now we need to grab the lines from the index.html page
#read in a HTML file
indexpage <- readr::read_lines("site/_site/index.html")

#store the string we want to find
findcode <- '<p>..</p>'
#create a vector to store the matches
linenums <- c()

for(linenum in seq(2,length(indexpage))){
  #if the line matches our target, store it
  if (indexpage[linenum] == findcode){
    print("HIT!")
    print(linenum)
    #store the line number - this will end up being the last one matched
    linenums <- c(linenums,linenum)
    }
}

linenums
```
Now we store the embed code that we want to put in those positions.

First we create four strings, each with a different embed code.

Then we store those in a vector.

Finally we loop through the vector of placeholders and replace each one with the embed code at the corresponding position.

```{r insert embed code}
embedcode1 <- '<div class="flourish-embed flourish-table" data-src="visualisation/14983909"><script src="https://public.flourish.studio/resources/embed.js"></script></div>'

#store the embed code from Flourish
embedcode2 <- '<div class="flourish-embed flourish-chart" data-src="visualisation/15389778"><script src="https://public.flourish.studio/resources/embed.js"></script></div>'

embedcode3 <- '<div class="flourish-embed flourish-table" data-src="visualisation/14350352"><script src="https://public.flourish.studio/resources/embed.js"></script></div>'

embedcode4 <- '<div class="flourish-embed flourish-hierarchy" data-src="visualisation/14992979"><script src="https://public.flourish.studio/resources/embed.js"></script></div>'

embedcode5 <- '<div class="flourish-embed flourish-hierarchy" data-src="visualisation/15307268"><script src="https://public.flourish.studio/resources/embed.js"></script></div>'

embedcode6 <- '<div class="flourish-embed flourish-hierarchy" data-src="visualisation/15310922"><script src="https://public.flourish.studio/resources/embed.js"></script></div>'

#store those codes in a vector
embedlist <- c(embedcode1, embedcode2, embedcode3, embedcode4, embedcode5, embedcode6)

#loop through a sequence of indices
for (i in seq(1,length(linenums))){
  print(linenums[i])
  indexpage[linenums[i]] <- embedlist[i] # replace the line number at that position with the embed code at that position
  }


#write the results over the original
write(x = indexpage, file="site/_site/index.html")
```
## Change the title tag

At the moment the `<title>` tag of the home page shows the name of the trust that was used for testing. We need to change that.

```{r change title tag}
indexpage[14] <- "<title>Analysing hospital discharge delays</title>" # replace the line number at that position with the text

#write the results over the original
write(x = indexpage, file="site/_site/index.html")
```

